---
title: "Data Preprocessing"
output:
  html_document:
    toc: yes
    toc_float:
      toc_collapsed: yes
    toc_depth: 5
    code_folding: hide
---


<h1><strong>DATA PREPROCESSING</strong></h1>

&nbsp;

In this section, we would like to introduce the data sources we used in our project, including the variable overview, and provide the procedure of data preprocessing.

&nbsp;

&nbsp;

## Data Sources


### DOHMH Rodent Inspection Data


The [rodent inspection data](https://data.cityofnewyork.us/Health/Rodent-Inspection/p937-wjvj) contains the information on rat sightings and intervention visits in NYC from 1918 to 2022, which is managed by New York City Department of Health and Mental Hygiene, Division of Environmental Health. The data source is the Veterinary, Rodent and Vector Surveillance System (VRVSS). It is also available on the [Rat Information Portal](https://www.nyc.gov/site/doh/health/health-topics/rats.page), which is a web-based mapping application where users can view and map rat inspection and intervention data. Users can search results from Health Department inspections for rats at the level of individual properties, and view neighborhood maps. 

What should to be mentioned is that most of the inspection is due to the complaints from the general public. Thus, if a property/taxlot does not appear in the file, which does not indicate an absence of rats - rather just that it has not been inspected. Similarly, neighborhoods with higher numbers properties with active rat signs may not actually have higher rat populations but simply have more inspections.


&nbsp;


### NYC Daily Weather Data


The [NYC daily weather data](https://www.ncei.noaa.gov/metadata/geoportal/rest/metadata/item/gov.noaa.ncdc:C00861/html) from the National Oceanic and Atmospheric Association (NOAA) of the National Centers for Environmental Information (NCEI), consists of the NYC daily climate observations from the Central Park station (id: USW00094728), which is integrated by the GHCN (Global Historical Climatology Network)-Daily database. The data is also available to be retrieved using R functions from the `rnoaa` package. 

Global Historical Climate Network includes daily land surface observations from over 100,000 stations in 180 countries and territories. The GHCN-Daily was developed to meet the needs of climate analysis and monitoring studies that require data at a sub-monthly time resolution (e.g., assessments of the frequency of heavy rainfall, heat wave duration, etc.). NCEI provides numerous daily variables, including maximum and minimum temperature, total daily precipitation, snowfall, and snow depth; however, about one half of the stations report precipitation only. Both the record length and period of record vary by station and cover intervals ranging from less than a year to more than 175 years. 


&nbsp;


### NYC COVID-19 Daily Case Count Data


The [NYC COVID-19 daily case count data](https://data.cityofnewyork.us/Health/COVID-19-Daily-Counts-of-Cases-Hospitalizations-an/rc75-m7u3), provided by New York City Department of Health and Mental Hygiene (DOHMH), represents citywide and borough-specific daily counts of COVID-19 confirmed cases and COVID-related hospitalizations and confirmed and probable deaths among New York City residents. The case counts are aggregated by date of diagnosis (i.e., date of specimen collection), and the hospitalization cases areaggregated by date of admission, and death cases are aggregated by date of death. 

Data collection is since February 29, 2020, which is the date that the Health Department classifies as the start of the COVID-19 outbreak in NYC as it was the date of the first laboratory-confirmed COVID-19 case. Data on confirmed cases were passively reported to the NYC Health Department by hospital, commercial, and public health laboratories. In March, April, and early May, the NYC Health Department had discouraged people with mild and moderate symptoms from being tested, so our data primarily represent people with more severe illness. Until mid-May, patients with more severe COVID-19 illness were more likely to have been tested and included in these data. Data on hospitalizations for confirmed COVID-19 cases were obtained from direct remote access to electronic health record systems, regional health information organization (RHIOs), and NYC Health and Hospital information, as well as matching to syndromic surveillance. Deaths were confirmed by the New York City Office of the Chief Medical Examiner and the Health Department’s Bureau of Vital Statistics. Data counts are underestimates. This dataset has been available to the public since May 19, 2020, and is updated on a daily basis.

These data can be used to:  

* Identify temporal trends in the number of persons diagnosed with COVID-19, citywide and by borough.
* Identify temporal trends in the numbers of COVID-19-related hospitalizations and deaths, citywide and by borough.

&nbsp;


### NYC Borough Population Data


The [NYC borough population data](http://www.citypopulation.de/en/usa/newyorkcity/), provided by City Population, consists of the population of the boroughs of New York City according to the U.S. Census Bureau results. Additional information about the population structure including gender, age groups, age distribution, race, ethnicity can be also found.


&nbsp;

&nbsp;

## Data Cleaning


Due to the large size of our final project datasets, another [repository](https://github.com/YijiaJiang/p8105_final_project_data) is created for storing tidied data. The codes below describes the steps we took to collate and merge dataset we used in the following exploratory and statistical analysis. Moreover, the detailed data preprocessing procedures can be find [here](https://github.com/YijiaJiang/p8105_final_project_data/blob/main/data_preprocessing_rat.Rmd). 


```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(tidyverse)
library(RSocrata)
library(rnoaa)
```

```{r}
# deal with rat inspection data
url_rat = "https://data.cityofnewyork.us/OData.svc/p937-wjvj"
rat = read.socrata(url_rat) %>% 
  janitor::clean_names()

rat_tidy = rat %>%
  select(inspection_type, bbl, zip_code, street_name, latitude, longitude, borough, result, inspection_date, approved_date) %>%
  drop_na() %>% 
  mutate(boro_code = substr(bbl, 1, 1),
         block = substr(bbl, 2, 6),
         lot = substr(bbl, 7, 10)) %>% 
  select(inspection_type, boro_code, block, lot, zip_code, street_name, latitude, longitude, borough, result, inspection_date, approved_date) %>% 
  separate(inspection_date, c("inspection_date", "inspection_time"), " ") %>% 
  separate(inspection_date, c("inspection_year", "inspection_month", "inspection_day"), "-") %>%
  separate(approved_date, c("approved_date", "approved_time"), " ") %>% 
  separate(approved_date, c("approved_year", "approved_month", "approved_day"), "-") %>%
  mutate(inspection_year = as.integer(inspection_year), 
         inspection_month = as.integer(inspection_month), 
         inspection_day = as.integer(inspection_day)) %>%
  mutate(approved_year = as.integer(approved_year), 
         approved_month = as.integer(approved_month), 
         approved_day = as.integer(approved_day)) %>%
  relocate(inspection_year, .before = "inspection_month") %>% 
  relocate(approved_year, .before = "approved_month") %>%
  arrange(inspection_year, inspection_month) %>% 
  mutate(inspection_month = month.abb[inspection_month],
         approved_month = month.abb[approved_month]) %>% 
  filter(inspection_year >= 2012 & inspection_year <= 2021)


# deal with weather data
nycstationsid = ghcnd_stations() %>% 
  filter(id == "USW00094728") %>% 
  distinct(id)

nyc_weather = meteo_pull_monitors(nycstationsid$id, 
                             date_min = "2012-01-01", 
                             date_max = "2021-12-31",
                             var = c("PRCP", "SNOW", "SNWD", "TMAX", "TMIN"))

nyc_weather_tidy = nyc_weather %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year),
         month = month.abb[as.numeric(month)],
         day = as.numeric(day)) %>%
  mutate(prcp = prcp/10,
         tmax = tmax/10,
         tmin = tmin/10) 


# deal with covid data
url_covid = "https://data.cityofnewyork.us/OData.svc/rc75-m7u3"
covid = read.socrata(url_covid) %>% 
  janitor::clean_names() 

covid_tidy = covid %>%
  rename(date = date_of_interest) %>% 
  select(date, contains("case_count")) %>% 
  select(-contains(c("probable_case_count", "case_count_7day_avg", "all_case_count_7day_avg"))) %>%
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year),
         month = month.abb[as.numeric(month)],
         day = as.numeric(day)) %>%
  pivot_longer(
    cols = bx_case_count:si_case_count,
    names_to = "borough",
    values_to = "borough_case_count"
  ) %>% 
  mutate(borough = gsub("_case_count", "", borough)) %>% 
  mutate(borough = dplyr::recode(borough, "bx" = "Bronx","bk" = "Brooklyn","mn" = "Manhattan","si" = "Staten Island","qn" = "Queens")) %>% 
  relocate(case_count, .after = borough_case_count) %>% 
  rename(total_case_count = case_count) 


# merge the above dataframe with covid info.
rat_weather_covid = rat_tidy %>% 
  filter(inspection_year %in% c(2020,2021)) %>% 
  merge(nyc_weather_tidy, by.x = c("inspection_year","inspection_month","inspection_day"), by.y = c("year","month","day")) %>% 
  select(-id) %>% 
  merge(covid_tidy, by.x = c("inspection_year","inspection_month","inspection_day","borough"), by.y = c("year","month","day","borough"))
```

&nbsp;

After cleaning the datasets mentioned above, our tidied and aggregated data has a total of 1,679,675 rows, one for each rat inspection record, and is the basis of our exploratory and statistical analysis. Altogether 24 variables are selected as meaningful and valuable. The specific variable names and their corresponding explanation are listed below:

* `inspection_year`: year of the inspection
* `inspection_month`: month of the inspection 
* `inspection_day`: day of the inspection
* `boro_code`: code assigned to the NYC borough
* `block`: block number for the inspected taxlot; block numbers repeat in different boroughs
* `lot`: lot number for the inspected taxlot (Notes: lot numbers can repeat in different blocks)
* `zip_code`: postal zipcode of the taxlot that was inspected
* `street_name`:
* `latitude`: latitude in decimal degrees of the inspected taxlot (WGS 1984)
* `longitude`: longitude in decimal degrees of the inspected taxlot (WGS 1984)
* `borough`: name of the NYC borough
* `result`: result of the inspection including Active Rat Signs (ARS) and Problem Conditions
* `inspection_time`: time of the inspection.
* `approved_year`: year of the inspection approved by supervisor at DOHMH
* `approved_month`: month of the inspection approved by supervisor at DOHMH
* `approved_day`: day of the inspection approved by supervisor at DOHMH
* `approved_time`: time of the inspection approved by supervisor at DOHMH
* `prcp`: precipitation (mm)
* `snow`: snowfall (mm)
* `snwd`: snow depth (mm)
* `tmax`: maximum temperature (&deg;C)
* `tmin`: minimum temperature (&deg;C)
* `borough_case_count`: count of patients tested who were confirmed to be COVID-19 cases on date_of_interest in borough_of_interest
* `total_case_count`: total count of patients tested who were confirmed to be COVID-19 cases on date_of_interest in NYC





