---
title: "Regression Analysis"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 5
    code_folding: hide
---

<h1><strong>REGRESSION ANALYSIS</strong></h1>

&nbsp;

In this section, we aimed to construct a linear regression model and to generate the monthly number of rats observed using a variety of potential predictors, along with the model diagnostics and model validation. The regression model will be further utilized in the Shiny App for the rat activity prediction. The outcome was defined as the number of rat cases which were inspected in different borough per month.


&nbsp;

&nbsp;

```{r, include=FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(car)
library(ggplot2)
library(GGally)
library(rstatix)
library(modelr)
library(patchwork)
library(see)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r, include = FALSE, message = FALSE, warning = FALSE}
# import the rat dataset
url_2012 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2012.csv"
rat_2012 = read_csv(url(url_2012)) 
url_2013 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2013.csv"
rat_2013 = read_csv(url(url_2013)) 
url_2014 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2014.csv"
rat_2014 = read_csv(url(url_2014)) 
url_2015 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2015.csv"
rat_2015 = read_csv(url(url_2015)) 
url_2016 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2016.csv"
rat_2016 = read_csv(url(url_2016)) 
url_2017 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2017.csv"
rat_2017 = read_csv(url(url_2017)) 
url_2018 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2018.csv"
rat_2018 = read_csv(url(url_2018)) 
url_2019 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2019.csv"
rat_2019 = read_csv(url(url_2019)) 
url_2020 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2020.csv"
rat_2020 = read_csv(url(url_2020)) 
url_2021 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2021.csv"
rat_2021 = read_csv(url(url_2021)) 
```


```{r, include = FALSE, message = FALSE, warning = FALSE}
# tidy data
rat = bind_rows(rat_2012, rat_2013, rat_2014, rat_2015, rat_2016, rat_2017, rat_2018, rat_2019, rat_2020, rat_2021) %>%
  select(-...1) %>%
  mutate(inspection_month_n = case_when(inspection_month == "Jan" ~ 1,
                                        inspection_month == "Feb" ~ 2,
                                        inspection_month == "Mar" ~ 3,
                                        inspection_month == "Apr" ~ 4,
                                        inspection_month == "May" ~ 5,
                                        inspection_month == "Jun" ~ 6,
                                        inspection_month == "Jul" ~ 7,
                                        inspection_month == "Aug" ~ 8,
                                        inspection_month == "Sep" ~ 9,
                                        inspection_month == "Oct" ~ 10,
                                        inspection_month == "Nov" ~ 11,
                                        inspection_month == "Dec" ~ 12)) %>%
  mutate(date = paste(inspection_year, inspection_month_n, inspection_day, sep = "-")) %>%
  mutate(date = as.Date(date,format = "%Y-%m-%d")) 
```


## Exploratory Statistical Analyses

Before constructing the model, we performed the correlational analysis to identify relevant variables and their impacts. 

```{r, include = FALSE, message=FALSE, warning=FALSE}
rat_reg = rat %>%
  mutate(inspection_ym = paste(inspection_year, inspection_month_n, sep = "-")) %>%
  mutate(covid_yn = case_when(date < "2020-02-29" ~ 0,
                              TRUE ~ 1)) %>%
  mutate(covid_yn = as.factor(covid_yn)) %>%
  separate(inspection_time, into = c("hour", "minute", "second")) %>%
  mutate(hour = as.numeric(hour)) %>%
  mutate(inspection_daytime = case_when(
      hour < 6 | hour >= 18 ~ "Night",
      hour >= 6 & hour < 12 ~ "Morning",
      hour >= 12 & hour < 18 ~ "Afternoon"))

cases = rat_reg %>% 
  select(inspection_ym, borough, inspection_month_n) %>% 
  group_by(inspection_ym, borough) %>% 
  add_count(borough, inspection_ym, name = "borough_monthly_cases") %>%
  distinct()

rat_tidy = rat_reg %>%
  dplyr::select(boro_code, borough, inspection_ym, inspection_month, covid_yn, prcp, snow, snwd, tmin, tmax) %>%
  mutate(borough = as.factor(borough),
         inspection_month = as.factor(inspection_month)) %>%
  group_by(inspection_ym, borough) %>% 
  summarise(avg_prcp = mean(prcp), 
            avg_snow = mean(snow), 
            avg_snwd = mean(snwd),
            avg_tmin = mean(tmin),
            avg_tmax = mean(tmax)) %>%
  mutate(avg_temp = (avg_tmin + avg_tmax)/2) %>% 
  mutate(covid_yn = case_when(inspection_ym < "2020-02" ~ 0, TRUE ~ 1)) %>% 
  mutate(boro_code = case_when(borough == "Manhattan" ~ 1,
                               borough == "Bronx" ~ 2,
                               borough == "Brooklyn" ~ 1,
                               borough == "Queens" ~ 4,
                               borough == "Staten Island" ~ 5))

cases_borough_monthly = merge(x = rat_tidy, y = cases, by = c("inspection_ym", "borough")) 
```



```{r, message=FALSE, warning=FALSE, fig.height=16, fig.width=16}
# correlation between key predictors
cases_borough_monthly %>%
  dplyr::select(borough, inspection_month_n, avg_prcp, avg_snow, avg_snwd, avg_tmin, avg_tmax, avg_temp, covid_yn) %>% 
  mutate(inspection_month_n = as.factor(inspection_month_n),
         covid_yn = as.factor(covid_yn)) %>% 
  rename(
    "Borough" = borough,
    "Month" = inspection_month_n,
    "Average Precipitation" = avg_prcp,
    "Average Snowfall" = avg_snow,
    "Average Snow Depth" = avg_snwd,
    "Average Tmin" = avg_tmin,
    "Average Tmax" = avg_tmax,
    "Average Temperature" = avg_temp,
    "COVID" = covid_yn
  ) %>% 
  ggpairs() + 
  scale_fill_discrete()
```

Through the correlation analysis between key predictors, the result above shown that the daily average temperature, the highest temperature, and the lowest temperature are highly correlated between each other.

&nbsp;

```{r, message=FALSE, warning=FALSE, fig.height=6, fig.width=6}
# correlation between predictors and outcome
cases_borough_monthly %>% 
  dplyr::select(boro_code, inspection_month_n, avg_prcp, avg_snow, avg_snwd, avg_tmin, avg_tmax, avg_temp, covid_yn, borough_monthly_cases) %>%
  rename(
    "Borough" = boro_code,
    "Month" = inspection_month_n,
    "Average Precipitation" = avg_prcp,
    "Average Snowfall" = avg_snow,
    "Average Snow Depth" = avg_snwd,
    "Average Tmin" = avg_tmin,
    "Average Tmax" = avg_tmax,
    "Average Temperature" = avg_temp,
    "COVID" = covid_yn
  ) %>% 
  cor_mat() %>%
  cor_gather() %>%
  filter(var1 %in% "borough_monthly_cases") %>%
  filter(!var2 %in% "borough_monthly_cases") %>%
  ggplot(aes(x = var1, y = var2, fill = cor, label = cor)) + 
  geom_tile(color = "white") +   
  geom_text(color = "white",size = 4) + 
  scale_x_discrete(labels = c("Borough Monthly cases")) + 
  labs(x = "Outcome Variable", y = "Predictor Variables",
       fill = "Correlation")
```


Through the correlation analysis between predictors and outcome, the result above shown that the borough was comparatively related to the outcome.

&nbsp;

&nbsp;


## Model Selection

### Full Model
```{r, message = FALSE, warning = FALSE}
# full model
cases_borough_monthly = cases_borough_monthly %>% 
  mutate(inspection_month = month.abb[inspection_month_n],
         covid_yn = as.factor(covid_yn))


model_linear_full = lm(borough_monthly_cases ~ inspection_month + borough + covid_yn + avg_prcp + avg_snow + avg_snwd + avg_temp, data = cases_borough_monthly) 

broom::tidy(model_linear_full) %>% 
   knitr::kable()

#summary(model_linear_full)
```

The full multiple linear regression model we constructed is 

$borough\_monthly\_cases= \beta_0+ \beta_1(inspectionMonth) + \beta_2(borough) + \beta_3(covid) + \beta_4(avg\_precipitaiton) + \\\beta_5(avg\_snowFall) + \beta_6(avg\_snowDepth) + \beta_7(avg\_temperature)$.   

The result shown that the p-value of `inspectionMonth`, `avg_precipitaiton`, `avg_snowFall`, and `avg_temperature` were greater than 0.05, which had to be analyzed by stepwise regression method.

&nbsp;

&nbsp;

### Stepwise Selection in Full Model

```{r, message = FALSE, warning = FALSE}
# stepwise with AIC (full model)
model_linear = step(model_linear_full, direction = "both")

#model_linear = lm(borough_monthly_cases ~ borough + covid_yn + avg_snow + avg_snwd, data = cases_borough_monthly) 

 
 broom::tidy(model_linear) %>% 
   knitr::kable()

#summary(model_linear)
```


We eliminated those variables that are not statistically significant using stepwise regression: 
$borough\_monthly\_cases= \beta_0+ \beta_1(borough) + \beta_2(covid) + \beta_3(snowFall) + \beta_4(snowDepth)$

&nbsp;

&nbsp;

### Full Model with Log Transformation

```{r, message = FALSE, warning = FALSE}
# full model with log transformation
model_linear_full_log = lm(log(borough_monthly_cases) ~ inspection_month + borough + covid_yn + avg_prcp + avg_snow + avg_snwd + avg_temp, data = cases_borough_monthly) 

 broom::tidy(model_linear_full_log) %>% 
   knitr::kable()

#summary(model_linear_full_log)
```

Log transformation was applied to our model to improve validity, additivity, and linearity:
$log(borough\_monthly\_cases)= \beta_0+ \beta_1(inspectionMonth) + \beta_2(borough) + \beta_3(covid) + \beta_4(avg\_precipitaiton) + \\\beta_5(avg\_snowFall) + \beta_6(avg\_snowDepth) + \beta_7(avg\_temperature)$

&nbsp;

&nbsp;

### Stepwise Selection in Full Model with Log Transformation

```{r, message = FALSE, warning = FALSE}
# stepwise with AIC
model_linear_log = step(model_linear_full_log, direction = "both")

#model_linear_log = lm(log(borough_monthly_cases) ~ inspection_month + borough + covid_yn + avg_prcp + avg_snwd + avg_temp, data = cases_borough_monthly) 


 broom::tidy(model_linear_log) %>% 
   knitr::kable()


#summary(model_linear_log)
```

Same as the original linear model above, we also applied the stepwise selection and got the final model:
$log(borough\_monthly\_cases)= \beta_0+ \beta_1(inspectionMonth) + \beta_2(borough) + \beta_3(covid) + \beta_4(avg\_precipitaiton)+ \\\beta_5(avg\_snowDepth) + \beta_6(avg\_temperature)$

&nbsp;

&nbsp;

## Model Diagnostics

```{r, message = FALSE, warning = FALSE, fig.width = 12,  fig.height = 9}
performance::check_model(model_linear, check = c("linearity", "qq", "normality", "outliers", "homogeneity", "vif"))
```


```{r, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 9}

performance::check_model(model_linear_log, check = c("linearity", "qq", "normality", "outliers", "homogeneity", "vif"))

```

According to plots above, we can see that linearity, homogeneity, and normality of residuals were all improved after the log transformation. Residuals were distributed around the 0 horizontal line and the reference line was flatter. There was no influential points outside the contour lines. Also, the residuals were approximately normally distributed.

&nbsp;

&nbsp;


## Cross-validation


```{r, message=FALSE, warning=FALSE, fig.height = 6}
# cross-validation 
set.seed(2022)

# without log
cv_df_1 = 
  crossv_mc(cases_borough_monthly, 100) %>% 
    mutate(
        train = map(train, as_tibble),
        test = map(test,as_tibble)
    )  %>%
  mutate(
    model_fit1  = map(train, ~lm(borough_monthly_cases ~ inspection_month + borough + covid_yn + avg_prcp + avg_snow + avg_snwd + avg_temp, data = .x)),
    model_fit2  = map(train, ~lm(borough_monthly_cases ~ borough + covid_yn + avg_snow + avg_snwd, data = .x))) %>% 
  mutate(
    rmse_1 = map2_dbl(model_fit1, test, ~rmse(model = .x, data = .y)),
    rmse_2 = map2_dbl(model_fit2, test, ~rmse(model = .x, data = .y))) 

p_cv_1 = cv_df_1 %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(title = "Comparison of the Cross-Validated Prediction Error", 
       x = "Models", 
       y = "Root Mean Square Error (RMSE)")  +
  scale_x_discrete(labels = c("Full Model", "Stepwise Model")) 

# log
cv_df_2 = 
  crossv_mc(cases_borough_monthly, 100) %>% 
    mutate(
        train = map(train, as_tibble),
        test = map(test,as_tibble)
    )  %>%
  mutate(
    model_fit3  = map(train, ~lm(log(borough_monthly_cases) ~ inspection_month + borough + covid_yn + avg_prcp + avg_snow + avg_snwd + avg_temp, data = .x)),
    model_fit4  = map(train, ~lm(log(borough_monthly_cases) ~ inspection_month + borough + covid_yn + avg_prcp + avg_snwd + avg_temp, data = .x))) %>% 
  mutate(
    rmse_3 = map2_dbl(model_fit3, test, ~rmse(model = .x, data = .y)),
    rmse_4 = map2_dbl(model_fit4, test, ~rmse(model = .x, data = .y))) 

p_cv_2 = cv_df_2 %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(title = "Comparison of the Cross-Validated Prediction Error", 
       x = "Models", 
       y = "Root Mean Square Error (RMSE)")  +
  scale_x_discrete(labels = c("Full Model (Log)", "Stepwise Model (Log)")) 


p_cv_1 / p_cv_2


```

According to the plot, there was a significant difference on RMSE between models with and without log-transformed outcomes. Model with log-transformation had a substantially lower average value of RMSE, which means this type of model better fits. After stepwise selection, only one variable was dropped and the RMSE were similar. We preferred the model with less variables to avoid potential over-fit.

&nbsp;

&nbsp;

## Summaries

The model diagnostics result shown linearity and normality were relatively improved after the log transformation and the stepwise regression. Therefore, we concluded that the number of rat inspection would be affected by  borough, COVID-19, and inspection time, also positively correlated with the precipitation and environment temperature, negatively related with the snow depth on the inspection date.


```{r, eval = FALSE, include=FALSE, message = FALSE, warning = FALSE}
## Logistic Regression Model
## We set the Result variable as the binary outcome. Passed(0) means none was inspected. 
rat_binary = rat %>% 
  dplyr::select(result, borough, inspection_month, inspection_year, inspection_daytime, covid_yn, prcp, snow, snwd, tmin, tmax) %>%
  mutate(inspection_result = case_when(
    result == "Passed" ~ 0, 
    result == "Rat Activity" ~ 1)) %>% 
  drop_na(inspection_result) %>%
  mutate(inspection_result = as.factor(inspection_result), 
         borough = as.factor(borough),
         covid_yn = as.factor(covid_yn),
         inspection_month = as.factor(inspection_month), 
         inspection_daytime = as.factor(inspection_daytime)) %>%
  mutate(avg_temp = (tmin + tmax)/2) %>% 
  mutate(feeling = as.character(avg_temp)) %>% 
  mutate(feeling = case_when(
      avg_temp <= 0 ~ 'Frozen',
      avg_temp <= 10 ~ 'Cold',
      avg_temp <= 20 ~ 'Cool',
      avg_temp <= 30 ~ 'Warm',
      avg_temp <= 40 ~ 'Hot')) %>% 
  mutate(prcp_yn = ifelse(prcp == 0,0,1),
         snow_yn = ifelse(snow == 0,0,1),
         snwd_yn = ifelse(snwd == 0,0,1)) %>% 
  mutate(across(.cols = c(prcp_yn, snow_yn, snwd_yn, feeling), as.factor)) %>% 
  filter(inspection_year == 2021) %>% 
  distinct()
```




```{r, eval = FALSE, include=FALSE, message = FALSE, warning = FALSE}
### Model Fitting
model_logit_full = glm(inspection_result ~ borough + inspection_daytime + covid_yn + prcp_yn + snow_yn + snwd_yn + feeling, data = rat_binary, family = "binomial")

broom::tidy(model_logit_full) %>% 
  knitr::kable()

summary(model_logit_full)
```


```{r, eval = FALSE, include=FALSE, message = FALSE, warning = FALSE}
### Model Diagnostics
set.seed(100)
par(mfrow = c(2,2))
plot(model_logit)
anova(model_logit)
```

