---
title: "Regression"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 5
    code_folding: hide
---

<h1><strong>REGRESSION</strong></h1>

In this section, we would like to construct a linear model to, together with a logistic model to predict. These two models will be further utilized in the Shiny App for rat activity prediction.

&nbsp;

&nbsp;

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(car)
```


```{r, message = FALSE, warning = FALSE}
# import dataset
url_2012 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2012.csv"
rat_2012 = read_csv(url(url_2012)) 
url_2013 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2013.csv"
rat_2013 = read_csv(url(url_2013)) 
url_2014 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2014.csv"
rat_2014 = read_csv(url(url_2014)) 
url_2015 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2015.csv"
rat_2015 = read_csv(url(url_2015)) 
url_2016 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2016.csv"
rat_2016 = read_csv(url(url_2016)) 
url_2017 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2017.csv"
rat_2017 = read_csv(url(url_2017)) 
url_2018 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2018.csv"
rat_2018 = read_csv(url(url_2018)) 
url_2019 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2019.csv"
rat_2019 = read_csv(url(url_2019)) 
url_2020 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2020.csv"
rat_2020 = read_csv(url(url_2020)) 
url_2021 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2021.csv"
rat_2021 = read_csv(url(url_2021)) 
```


```{r, message = FALSE, warning = FALSE}
rat = bind_rows(rat_2012, rat_2013, rat_2014, rat_2015, rat_2016, rat_2017, rat_2018, rat_2019, rat_2020, rat_2021) %>%
  select(-...1) %>%
   mutate(inspection_month_n = case_when(inspection_month == "Jan" ~ 1,
                                         inspection_month == "Feb" ~ 2,
                                         inspection_month == "Mar" ~ 3,
                                         inspection_month == "Apr" ~ 4,
                                         inspection_month == "May" ~ 5,
                                         inspection_month == "Jun" ~ 6,
                                         inspection_month == "Jul" ~ 7,
                                         inspection_month == "Aug" ~ 8,
                                         inspection_month == "Sep" ~ 9,
                                         inspection_month == "Oct" ~ 10,
                                         inspection_month == "Nov" ~ 11,
                                         inspection_month == "Dec" ~ 12)) %>%
  mutate(date = paste(inspection_year, inspection_month_n, inspection_day, sep = "-")) %>%
  mutate(date = as.Date(date,format = "%Y-%m-%d")) %>% 
  mutate(inspection_ym = paste(inspection_year, inspection_month_n, sep = "-")) %>%
  mutate(covid_yn = case_when(date < "2020-02-29" ~ 0,
                              TRUE ~ 1)) %>%
  separate(inspection_time, into = c("hour", "minute", "second")) %>%
  mutate(hour = as.numeric(hour)) %>%
  mutate(inspection_daytime = case_when(
      hour < 6 | hour >= 18 ~ "Night",
      hour >= 6 & hour < 12 ~ "Morning",
      hour >= 12 & hour < 18 ~ "Afternoon"))
  
#%>%  filter(result != "Passed")


```



## Exploratory Statistical Analyses

```{r, message=FALSE, warning=FALSE, fig.height=8, fig.width=8}
cases = rat %>% 
  mutate(inspection_ym = paste(inspection_year, inspection_month_n)) %>%
  filter(result != "Passed") %>%
  select(inspection_ym, borough) %>% 
  group_by(inspection_ym, borough) %>% 
  add_count(borough, inspection_ym, name = "borough_month_cases") %>%
  distinct()

rat_tidy = rat %>% 
  mutate(inspection_ym = paste(inspection_year, inspection_month_n)) %>%
  filter(result != "Passed") %>%
  select(boro_code, borough, inspection_ym, inspection_month, inspection_month_n, prcp, snow, snwd, tmin, tmax) %>%
  mutate(borough = as.factor(borough), inspection_month = as.factor(inspection_month), 
         prcp = as.numeric(prcp), snow = as.numeric(snow), snwd = as.numeric(snwd),
         tmin = as.numeric(tmin), tmax = as.numeric(tmax)) 

cases_borough_month = merge(x = rat_tidy, y = cases, by = c("inspection_ym", "borough")) %>% distinct() 


cases_borough_month %>%
  select(borough, inspection_month, prcp, snow, snwd, tmin, tmax) %>% 
  rename(
    "Borough" = borough,
    "Month" = inspection_month,
    "Precipitation" = prcp,
    "Snowfall" = snow,
    "Snow Depth" = snwd,
    "Minimum Temperature" = tmin,
    "Maximum Temperature" = tmax
  ) %>% 
  ggpairs(
    title = "Correlations Between Key Outcomes"
  ) + 
  scale_fill_discrete()
```


```{r, message=FALSE, warning=FALSE}
cases_borough_month %>%
  select(-inspection_ym, -borough) %>%
  rename(
    "Borough" = boro_code,
    "Month" = inspection_month,
    "Precipitation" = prcp,
    "Snowfall" = snow,
    "Snow Depth" = snwd,
    "Minimum Temperature" = tmin,
    "Maximum Temperature" = tmax,
    "Borough Monthly Cases" = borough_month_cases
  ) %>% 
  ggcorr(label = TRUE, hjust = 0.9, layout.exp = 2, label_size = 3, label_round = 2)


cases_borough_month %>% 
  select(boro_code, inspection_month_n, tmin, tmax, prcp, snow, snwd, borough_month_cases) %>%
  cor_mat() %>%
  cor_gather() %>%
  filter(var1 %in% "borough_month_cases") %>%
  filter(!var2 %in% "borough_month_cases") %>%
  mutate(
    sig_p = ifelse(p < 0.01, T, F),
    cor_if_sig = ifelse(p < 0.01, cor, NA)
    ) %>% 
  ggplot(aes(
    x = var1, 
    y = var2, 
    fill = cor,
    label = round(cor_if_sig, 2))) + 
  geom_tile(color = "white") +   
  geom_text(
    color = "white",
    size = 4
  ) + 
  scale_x_discrete(
    labels = c("Borough Monthly cases")
  ) + 
  labs(
    x = "Outcome Variable",
    y = "Predictor Variables",
    title = "Correlation Matrix between Predictors and Outcome",
    subtitle = "significant predictors at significance level 0.01",
    fill = "Correlation"
  )


skimr::skim(cases_borough_month)
```



## Variable selection
 
```{r, message = FALSE, warning = FALSE,comment = FALSE}
cases = rat %>% 
  select(borough, inspection_ym, inspection_month, covid_yn, prcp, snow, snwd, tmin, tmax) %>% 
  group_by(inspection_ym, borough) %>% 
  add_count(borough, inspection_ym, name = "borough_monthly_cases") %>%
  select(-prcp, -snow, -snwd, -tmin, -tmax) %>% distinct()

# 
 rat_tidy = rat %>%
   select(borough, inspection_ym, inspection_month, covid_yn, prcp, snow, snwd, tmin, tmax) %>%
   mutate(borough = as.factor(borough),
          inspection_month = as.factor(inspection_month),
          covid_yn = as.factor(covid_yn)) %>%
   group_by(inspection_ym, borough) %>% 
   summarise(avg_prcp = mean(prcp), 
             avg_snow = mean(snow), 
             avg_snwd = mean(snwd),
             avg_tmin = mean(tmin),
             avg_tmax = mean(tmax)) %>%
  mutate(avg_temp = (avg_tmin + avg_tmax)/2) %>% 
  mutate(Feeling = as.character(avg_temp)) %>% 
  mutate(Feeling = case_when(
      avg_temp <= 0 ~ 'Frozen',
      avg_temp <= 10 ~ 'Cold',
      avg_temp <= 20 ~ 'Cool',
      avg_temp <= 30 ~ 'Warm',
      avg_temp <= 40 ~ 'Hot')) 


cases_borough_monthly = merge(x = rat_tidy, y = cases, by = c("inspection_ym", "borough")) %>% distinct() 
```

## Linear Regression Model

```{r, message = FALSE, warning = FALSE,comment = FALSE}
  
model_linear_original = lm(borough_monthly_cases ~ inspection_month + borough + covid_yn + avg_prcp + avg_snwd + Feeling, data = cases_borough_monthly) 
broom::tidy(model_linear_original) %>% 
  knitr::kable()
summary(model_linear_original)

#model_linear = step(model_linear_original, direction = "backward")
#broom::tidy(model_linear) %>% 
#  knitr::kable()


#summary(model_linear)
```

## Model diagnostics

```{r, message = FALSE, warning = FALSE,comment = FALSE}
set.seed(100)
par(mfrow=c(2,2))
plot(model_linear_original)
anova(model_linear_original)
```


```{r, message = FALSE, warning = FALSE}

outlierTest(model_linear_original)

durbinWatsonTest(model_linear_original)

vif(model_linear_original)
```



## Logistic Regression Model

We set the Result variable as the binary outcome. Passed(0) means none was inspected.

```{r, message = FALSE, warning = FALSE,comment = FALSE}

rat_binary = rat %>% 
  select(result, borough, inspection_month, inspection_daytime, covid_yn, prcp, snow, snwd, tmin, tmax) %>%
  mutate(inspection_result = case_when(
    result == "Passed" ~ 0, 
    result == "Rat Activity" ~ 1)) %>% drop_na(inspection_result) %>%
  mutate(inspection_result = as.factor(inspection_result), 
         borough = as.factor(borough),
         covid_yn = as.factor(covid_yn),
         inspection_month = as.factor(inspection_month), 
         inspection_daytime = as.factor(inspection_daytime)) %>%
  mutate(Average_Temperature = (tmin + tmax)/2) %>% 
  mutate(Feeling = as.character(Average_Temperature)) %>% 
  mutate(Feeling = case_when(
      Average_Temperature <= 0 ~ 'Frozen',
      Average_Temperature <= 10 ~ 'Cold',
      Average_Temperature <= 20 ~ 'Cool',
      Average_Temperature <= 30 ~ 'Warm',
      Average_Temperature <= 40 ~ 'Hot')) %>% 
  distinct()



```



```{r, message = FALSE, warning = FALSE}
model_logit_original = glm(inspection_result ~ inspection_month + borough + covid_yn + inspection_daytime + prcp + snow + snwd + Feeling, data = rat_binary, family="binomial")

model_logit = step(model_logit_original, direction = "backward")
#broom::tidy(model_logit) %>% 
#  knitr::kable()

summary(model_logit_original)
anova(model_logit_original)
```


```{r, message = FALSE, warning = FALSE,comment = FALSE}
set.seed(100)
par(mfrow=c(2,2))
plot(model_logit)
anova(model_logit)

```
## Merge with Covid-19


```{r, message = FALSE, warning = FALSE}

url_covid = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_weather_covid.csv"
rat_covid = read_csv(url_covid)


rat_covid = rat_covid %>%
  mutate(inspection_month_n = case_when(inspection_month == "Jan" ~ 1,
                                   inspection_month == "Feb" ~ 2,
                                   inspection_month == "Mar" ~ 3,
                                   inspection_month == "Apr" ~ 4,
                                   inspection_month == "May" ~ 5,
                                   inspection_month == "Jun" ~ 6,
                                   inspection_month == "Jul" ~ 7,
                                   inspection_month == "Aug" ~ 8,
                                   inspection_month == "Sep" ~ 9,
                                   inspection_month == "Oct" ~ 10,
                                   inspection_month == "Nov" ~ 11,
                                   inspection_month == "Dec" ~ 12)) %>%
  mutate(date = paste(inspection_year, inspection_month_n, inspection_day, sep = "-"))%>%
  mutate(date = as.Date(date,format = "%Y-%m-%d"))

```
