---
title: "Statistical Analysis"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 5
    code_folding: hide
---

```{r, message = FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(patchwork)
library(plotly)
library(leaflet)
library(ggplot2)
library(lubridate)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r, message = FALSE, warning = FALSE, echo=FALSE}
url_2012 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2012.csv"
rat_2012 = read_csv(url(url_2012)) 
url_2013 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2013.csv"
rat_2013 = read_csv(url(url_2013)) 
url_2014 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2014.csv"
rat_2014 = read_csv(url(url_2014)) 
url_2015 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2015.csv"
rat_2015 = read_csv(url(url_2015)) 
url_2016 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2016.csv"
rat_2016 = read_csv(url(url_2016)) 
url_2017 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2017.csv"
rat_2017 = read_csv(url(url_2017)) 
url_2018 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2018.csv"
rat_2018 = read_csv(url(url_2018)) 
url_2019 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2019.csv"
rat_2019 = read_csv(url(url_2019)) 
url_2020 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2020.csv"
rat_2020 = read_csv(url(url_2020)) 
url_2021 = "https://raw.githubusercontent.com/YijiaJiang/p8105_final_project_data/main/rat_2021.csv"
rat_2021 = read_csv(url(url_2021)) 
rat = bind_rows(rat_2012, rat_2013, rat_2014, rat_2015, rat_2016, rat_2017, rat_2018, rat_2019, rat_2020, rat_2021) %>%
  select(-...1) %>%
   mutate(inspection_month_n = recode(inspection_month, 
                                   "Jan" = 1,
                                   "Feb" = 2,
                                   "Mar" = 3,
                                   "Apr" = 4,
                                   "May" = 5,
                                   "Jun" = 6,
                                   "Jul" = 7,
                                   "Aug" = 8,
                                   "Sep" = 9,
                                   "Oct" = 10,
                                   "Nov" = 11,
                                   "Dec" = 12)) %>%
  mutate(date = paste(inspection_year, inspection_month_n, inspection_day, sep = "-"))%>%
  mutate(date = as.Date(date,format = "%Y-%m-%d"))
```

## Boroughs

&nbsp;

### Chi-square testing homogenity among boroughs in recent years

We decide to look at the number of inspections over years to figure out the distribution of inspections varies in different boroughs from year to year.

$H_0$ : The distribution of inspections among boroughs are same over years.

$H_1$ : The distribution of inspections among boroughs aren’t all the same over years.

```{r, message = FALSE, warning = FALSE}
# create a data frame
rat_years = rat %>% 
  janitor::clean_names() %>% 
  dplyr::select(inspection_year, borough) %>% 
  group_by(borough, inspection_year) %>%
  summarize(frequency = n()) %>%
  mutate(borough = as.factor(borough),
    borough = fct_reorder(borough, frequency, .desc = TRUE)) %>% 
  pivot_wider(names_from = "inspection_year", values_from = "frequency")

# print the table
rat_years %>% t() %>% 
knitr::kable(caption = "Distribution of inspections across the years")

# Chi-square test
chisq.test(rat_years[,-1]) %>% 
   broom::tidy()
```

We can see from the Chi-square test above, we reject the null hypothesis at 99% significant level. This means that the inspections among boroughs have a different distribution across years.

&nbsp;

&nbsp;

### Chi-square testing homogenity among boroughs in 2021

Given that the distribution of inspections varies in different boroughs from year to year, we decide to look at inspections among different boroughs in 2021. 

$H_0$ : The distribution of inspections among boroughs are same in 2021.

$H_1$ : The distribution of inspections among boroughs aren’t all the same in 2021.

```{r, message = FALSE, warning = FALSE}
# import the data
rat_boro = rat_2021 %>% 
  janitor::clean_names() %>% 
  dplyr::select(inspection_year, borough) %>% 
  group_by(borough, inspection_year) %>% 
  summarize(frequency = n()) %>%
  mutate(borough = as.factor(borough),
    borough = fct_reorder(borough, frequency, .desc = TRUE)) %>% 
  dplyr::select(borough, frequency)

# print the table
knitr::kable(rat_boro, caption = "Distribution of inspections in 2021")

# chi-square test
chisq.test(rat_boro[,-1]) %>% 
  broom::tidy()
```

We can see from the chi-square test result we should reject the null hypothesis, the inspections are not equally distributed among boroughs in 2021.

&nbsp;

&nbsp;

## Weather and covid-19

&nbsp;

### Chi-sqaure testing association betwwen inspection density and temperature in 2021

$H_0$ : The distribution of inspections in different weather are same in 2021.

$H_1$ : The distribution of inspections in different weather aren’t all the same in 2021.

```{r, message = FALSE, warning = FALSE}
# create a data frame
rat_n = rat_2021 %>%
  group_by(inspection_year, inspection_month, inspection_day)

rat_t = rat %>%
  mutate(Average_Temperature = (tmin + tmax)/2) %>% 
  rename(Date = date) %>%
  mutate(Feeling = as.character(Average_Temperature)) %>% 
  mutate(Feeling = case_when(
      Average_Temperature <= 0 ~ 'Frozen',
      Average_Temperature <= 10 ~ 'Cold',
      Average_Temperature <= 20 ~ 'Cool',
      Average_Temperature <= 30 ~ 'Warm',
      Average_Temperature <= 40 ~ 'Hot')) %>% 
  distinct(inspection_year, inspection_month, inspection_day, Date, Average_Temperature, tmax, tmin, Feeling)

rat_temp = 
  left_join(rat_n, rat_t, by = c("inspection_year", "inspection_month", "inspection_day")) %>%
  dplyr::select(Average_Temperature, Feeling) %>% 
  group_by(Feeling) %>%
  summarize(frequency = n()) %>%
  mutate(Feeling = as.factor(Feeling),
    Feeling = fct_reorder(Feeling, frequency, .desc = TRUE)) %>% 
  pivot_wider(names_from = "Feeling", values_from = "frequency") 

# print the table
knitr::kable(rat_temp, caption = "The association between number of rat inspections and temperature in 2021")

# chi-square test
chisq.test(rat_temp[,-1]) %>% 
  broom::tidy()
```

We can see from the chi-square test result we should reject the null hypothesis, the inspections are not equally distributed in different weather in 2021.

&nbsp;

&nbsp;

### Wilcoxon Rank-sum testing changes of inspection density caused by covid-19

For the rodent inspection across time, we compared mean difference between inspection density in 2019 before covid and inspection density in 2020 after covid.

$H_0$ : The distribution of annual inspection before covid and after covid is same.

$H_1$ : The distribution of annual inspection before covid and after covid is different.

```{r, message = FALSE, warning = FALSE}
# import the data
rat_covid = rbind(rat_2019, rat_2020) %>% 
  janitor::clean_names() %>% 
  dplyr::select(inspection_year, inspection_month) %>% 
  mutate(Inspection_year = as.factor(inspection_year),
         Inspection_month = as.factor(inspection_month)) %>%
  mutate(Inspection_month = inspection_month %>% 
                       fct_relevel("Jan", "Feb", "Mar","Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) %>%
  group_by(Inspection_year,Inspection_month) %>%
  summarize(Frequency = n())

# print the table
rat_covid2 = rat_covid %>%   
  pivot_wider(names_from = "Inspection_year", values_from = "Frequency")
knitr::kable(rat_covid2, caption = "Number of rat inspections before and after covid")

# Wilcoxon Rank-sum test
wilcox.test(Frequency ~ Inspection_year, data = rat_covid,
                   exact = FALSE)
```

We can see from the Wilcoxon Rank-sum test result that the true location shift is not equal to 0, which means that the medians of two populations differ. Therefore, we have the evidence to reject the null hypothesis and conclude that there is a difference between inspection density in 2019 before covid and inspection density in 2020 after covid.
